---
title: "NBA19-20"
author: "Jaehack"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-"
)
```

1) Loading the season data and extracting some meaningful statistics 
 - field goals made
 - three points made
 - free throws made
 - total rebounds
 - total points
 
```{r}
library(SportsAnalytics)

nba1920 <- fetch_NBAPlayerStatistics("19-20")

player_stat <- nba1920[, c("Name", "Team", "FieldGoalsMade", "ThreesMade", "FreeThrowsMade", "TotalRebounds",
                           "TotalPoints")]
```

2) Subsetting the data to include only players for LAL, the champion of 19-20 season and finding the players who are “best” for some meaningful stats.

```{r}
LAL_subset <- subset(nba1920, Team == 'LAL')
sub <- matrix(c(LAL_subset[which.max(LAL_subset$FieldGoalsMade), c("Name", "FieldGoalsMade")],
LAL_subset[which.max(LAL_subset$ThreesMade), c("Name", "ThreesMade")],
LAL_subset[which.max(LAL_subset$FreeThrowsMade), c("Name", "FreeThrowsMade")],
LAL_subset[which.max(LAL_subset$TotalRebounds), c("Name", "TotalRebounds")],
LAL_subset[which.max(LAL_subset$TotalPoints), c("Name", "TotalPoints")]), ncol = 2, byrow = TRUE)
colnames(sub) <- c("Player", "Statistics")
rownames(sub) <- c("Field Goals Made", "Three Points Made", "Free Throws Made", "Total Rebounds",
                   "Total Points")
sub
```

3) Showing 5 teams for the season that have the most wins in descending order.

```{r}
library(rvest)
reg_season1920 <- read_html("https://en.wikipedia.org/wiki/2019%E2%80%9320_NBA_season")
reg_season_data <- reg_season1920 %>% html_nodes("table.wikitable")

#By Division
East_ATL_Div <- (reg_season_data[2] %>% html_table)[[1]]  # Eastern Conference - Atlantic Division
East_CEN_Div <- (reg_season_data[3] %>% html_table)[[1]]  # Eastern Conference - Central Division
East_SOU_Div <- (reg_season_data[4] %>% html_table)[[1]]  # Eastern Conference - Southeast Division

West_NOR_Div <- (reg_season_data[5] %>% html_table)[[1]]  # Western Conference - Northwest Division
West_PAC_Div <- (reg_season_data[6] %>% html_table)[[1]]  # Western Conference - Pacific Division
West_SOU_Div <- (reg_season_data[7] %>% html_table)[[1]]  # Western Conference - Southhwest Division

#Getting rid of unnecessary indications on the Team names
East_ATL_Div$`Atlantic Division` <- gsub("y – ", "", East_ATL_Div$`Atlantic Division`)
East_ATL_Div$`Atlantic Division` <- gsub("x – ", "", East_ATL_Div$`Atlantic Division`)

East_CEN_Div$`Central Division` <- gsub("z – ", "", East_CEN_Div$`Central Division`)
East_CEN_Div$`Central Division` <- gsub("x – ", "", East_CEN_Div$`Central Division`)

East_SOU_Div$`Southeast Division` <- gsub("y – ", "", East_SOU_Div$`Southeast Division`)
East_SOU_Div$`Southeast Division` <- gsub("x – ", "", East_SOU_Div$`Southeast Division`)

West_NOR_Div$`Northwest Division` <- gsub("y – ", "", West_NOR_Div$`Northwest Division`)
West_NOR_Div$`Northwest Division` <- gsub("x – ", "", West_NOR_Div$`Northwest Division`)

West_PAC_Div$`Pacific Division` <- gsub("c – ", "", West_PAC_Div$`Pacific Division`)
West_PAC_Div$`Pacific Division` <- gsub("x – ", "", West_PAC_Div$`Pacific Division`)

West_SOU_Div$`Southwest Division` <- gsub("y – ", "", West_SOU_Div$`Southwest Division`)
West_SOU_Div$`Southwest Division` <- gsub("x – ", "", West_SOU_Div$`Southwest Division`)

East_ATL_Div
East_CEN_Div
East_SOU_Div

West_NOR_Div
West_PAC_Div
West_SOU_Div

#By Conference with top 5 teams
EastConf <- (reg_season_data[8] %>% html_table)[[1]][1:6,]  #6 rows needed due to the header name (conference)
WestConf <- (reg_season_data[9] %>% html_table)[[1]][1:6,]

colnames(EastConf) <- EastConf[1,]  #Make the appropriate column names for the table
EastConf <- EastConf[-1,]
rownames(EastConf) <- NULL

colnames(WestConf) <- WestConf[1,]
WestConf <- WestConf[-1,]
rownames(WestConf) <- NULL

EastConf$Team <- gsub("z – ", "", EastConf$Team)
EastConf$Team <- gsub("y – ", "", EastConf$Team)
EastConf$Team <- gsub("x – ", "", EastConf$Team)
EastConf$Team <- gsub("\\*", "", EastConf$Team)
EastConf$Team <- gsub("\\s", "", EastConf$Team)
EastConf$Team <- gsub("([a-z])([A-Z])", "\\1 \\2", EastConf$Team)

WestConf$Team <- gsub("c – ", "", WestConf$Team)
WestConf$Team <- gsub("x – ", "", WestConf$Team)
WestConf$Team <- gsub("y – ", "", WestConf$Team)
WestConf$Team <- gsub("\\*", "", WestConf$Team)
WestConf$Team <- gsub("\\s", "", WestConf$Team)
WestConf$Team <- gsub("([a-z])([A-Z])", "\\1 \\2", WestConf$Team)

EastConf
WestConf
```

4) Creating data visualizations with NBA 19-29
i) Comparison between Eastern and Western in terms of five statistics that are previously selected

```{r}
library(ggplot2)
library(tidyr)

EastStat <- subset(player_stat, player_stat$Team == "MIL" | player_stat$Team == "TOR" | 
                     player_stat$Team == "BOS" | player_stat$Team == "IND" | player_stat$Team == "MIA" |
                     player_stat$Team == "PHI" | player_stat$Team == "BRO" | player_stat$Team == "ORL" |
                     player_stat$Team == "WAS" | player_stat$Team == "CHA" | player_stat$Team == "CHI" |
                     player_stat$Team == "NYK" | player_stat$Team == "DET" | player_stat$Team == "ATL" |
                     player_stat$Team == "CLE")
WestStat <- subset(player_stat, player_stat$Team == "LAL" | player_stat$Team == "LAC" | 
                     player_stat$Team == "DEN" | player_stat$Team == "HOU" | player_stat$Team == "OKL" |
                     player_stat$Team == "UTA" | player_stat$Team == "DAL" | player_stat$Team == "POR" |
                     player_stat$Team == "MEM" | player_stat$Team == "PHO" | player_stat$Team == "SAN" |
                     player_stat$Team == "SAC" | player_stat$Team == "NOR" | player_stat$Team == "MIN" |
                     player_stat$Team == "GSW")

EastStat$Conf <- "Eastern"
WestStat$Conf <- "Western"

AllConf <- rbind(EastStat, WestStat)

FGM_by_Conf <- setNames(aggregate(x = AllConf$FieldGoalsMade, by = list(AllConf$Conf), FUN = sum), 
                        c("Conference", "Field Goals Made"))
THR_by_Conf <- setNames(aggregate(x = AllConf$ThreesMade, by = list(AllConf$Conf), FUN = sum),
                        c("Conference", "Three Points Made"))
FTM_by_Conf <- setNames(aggregate(x = AllConf$FreeThrowsMade, by = list(AllConf$Conf), FUN = sum),
                        c("Conference", "Free Throws Made"))
TRB_by_Conf <- setNames(aggregate(x = AllConf$TotalRebounds, by = list(AllConf$Conf), FUN = sum),
                        c("Conference", "Total Rebounds"))
TP_by_Conf <- setNames(aggregate(x = AllConf$TotalPoints, by = list(AllConf$Conf), FUN = sum),
                       c("Conference", "Total Points"))

AllMerged <- merge(merge(merge(merge(FGM_by_Conf, THR_by_Conf), FTM_by_Conf), TRB_by_Conf), TP_by_Conf)

AllMergedGathered <- gather(AllMerged, Stats, Records, -Conference)
ggplot(AllMergedGathered, aes(x = Stats, y = Records, fill = factor(Conference))) + 
  geom_bar(stat = 'identity', position = 'dodge') + theme_minimal() + 
  scale_fill_discrete(name = "Conference") + 
  scale_fill_manual("Conference", values = c("Eastern" = "green", "Western" = "red")) + 
  theme(axis.text.x = element_text(face = "bold")) + theme(axis.text.y = element_text(face = "bold")) +
  ggtitle("Five Stats Comparison by Conference") + xlab('') + ylab('') +
  coord_flip()
```

The comparison above reflects every Eastern and Western conference team. As we can see, the Western statistics are slightly higher than the Eastern, but the difference does not seem to be meaningfully significant. Although it shows only the 19-20 season along with just five statistics, which are not enough to make a conclusion, as a general fan of NBA, I believe both conferences are the best basketball leagues in the world.

ii) Fouls by position of top 5 Western teams

```{r}
west_top5 <- subset(nba1920, nba1920$Team == "LAL" | nba1920$Team == "LAC" | nba1920$Team == "DEN" |
                      nba1920$Team == "HOU" | nba1920$Team == "OKL")

foul_by_pos <- west_top5[, c("Team", "Position", "PersonalFouls")]
ggplot(foul_by_pos, aes(x = Position, y = PersonalFouls, fill = Team)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  ggtitle("Personal Fouls by Position of Top 5 Western Teams") + 
  scale_fill_manual("Team", values = c("DEN" = "navy blue", "HOU" = "red", "LAC" = "black", 
                                       "LAL" = "yellow", "OKL" = "orange")) +
  ylab("Number of Fouls") + theme_bw() + theme(plot.title = element_text(face = "bold"))
```

The graph compares top 5 Western Conference teams in the season. Houston Rockets players committed many more fouls than other teams except Centers. By position, Centers and Power Forward similarly committed fouls except Houston Rockets, and LA Lakers Point Guards made the minimum fouls, while OKL City Thunder Small Forwards committed the minimum fouls. Lastly, the shooting guard of LA Clippers made the least fouls.

iii) Number of Views on NBA Final

```{r}
final_views <- read_html("https://en.wikipedia.org/wiki/NBA_Finals_television_ratings")
final_views_data <- final_views %>% html_nodes("table.wikitable")

views_by_year <- (final_views_data[2] %>% html_table(fill = TRUE))[[1]]
views_by_year <- views_by_year[, 1:ncol(views_by_year) - 1]

views_by_year$Year <- gsub('\\[a]', '', views_by_year$Year)

views_by_year$Avg <- gsub("[a-z]", "", views_by_year$Avg)
views_by_year$Avg <- gsub("[A-Z]", "", views_by_year$Avg)
views_by_year$Avg <- gsub(".*\\((.*)\\).*", "\\1", views_by_year$Avg)

views_by_year$Avg <- as.numeric(views_by_year$Avg)
views_by_year$Year <- as.numeric(views_by_year$Year)

views_by_year <- views_by_year[1:35, ] #Before 1986, there's no data of number of views, so truncated
views_by_year$NumGame <- ifelse(views_by_year$`Game 5` == "No Game" & !is.na(views_by_year$`Game 5`), 
                                "Sweep", views_by_year$`Game 5`)
views_by_year$NumGame <- ifelse(views_by_year$`Game 5` != "No Game" & views_by_year$`Game 6` == "No Game" & 
                                  !is.na(views_by_year$`Game 6`), "5 Games", views_by_year$NumGame)
views_by_year$NumGame <- ifelse(views_by_year$`Game 5` != "No Game" & views_by_year$`Game 6` != "No Game" & 
                                  views_by_year$`Game 7` == "No Game" & !is.na(views_by_year$`Game 7`), 
                                "6 Games", views_by_year$NumGame)
views_by_year$NumGame <- ifelse(views_by_year$`Game 5` != "No Game" & views_by_year$`Game 6` != "No Game" & 
                                  views_by_year$`Game 7` != "No Game" & !is.na(views_by_year$`Game 7`), 
                                "7 Games", views_by_year$NumGame)

library(gridExtra)
a <- ggplot(views_by_year, aes(x = Year, y = Avg)) + geom_line(linetype = 'dashed') + geom_point() + 
  ggtitle("Number of Views on NBA Finals since 1986") + ylab("Average Views in Million") + theme_bw() +
  theme(plot.title = element_text(face = "bold"))

b <-ggplot(views_by_year, aes(x = Year, y = Avg, group = NumGame, colour = NumGame)) + geom_line(size = 1) + 
  ggtitle("Number of Views on NBA Finals since 1986 by Games") + ylab("Average Views in Million") +
  theme_bw() + theme(plot.title = element_text(face = "bold"))

grid.arrange(a, b, nrow = 2)
```

Those graphs show the number of views on NBA Finals since 1986. Unfortunately, 2020 NBA final recorded the lowest views since 1986. NBA had a tough time in this year due to COVID-19, which seems to impact even the number of views on final games. Also, the number of views significantly dropped after 1990s. This seems to be due to the retirement of the greatest basket player, Michael Jordan (mostly retired in 90s as he retired several times). 

The second graph shows the number of views by number of final games. The sweep games are slightly less viewed than others, which means that the more final games, the higher likelihood of more views.

iv) Five NBA Star Players Stat Comparison

```{r}
star_players <- subset(nba1920, nba1920$Name == "Lebron James" | nba1920$Name == "G Antetokounmpo" |
                       nba1920$Name == "Anthony Davis" | nba1920$Name == "James Harden" | 
                       nba1920$Name == "Kawhi Leonard") 

TMP <- setNames(aggregate(x = star_players$TotalMinutesPlayed, by = list(star_players$Name), FUN = sum),
                c("Name", "Total Minutes Played"))
TP <- setNames(aggregate(x = star_players$TotalPoints, by = list(star_players$Name), FUN = sum),
               c("Name", "Total Points"))
TR <- setNames(aggregate(x = star_players$TotalRebounds, by = list(star_players$Name), FUN = sum),
                c("Name", "Total Rebound"))
FGM <- setNames(aggregate(x = star_players$FieldGoalsMade, by = list(star_players$Name), FUN = sum),
                c("Name", "Field Goals Made"))

AllMerged2 <- merge(merge(merge(TMP, TP), TR), FGM)

AllMerged2_gathered <- gather(AllMerged2, Stats, Records, -Name)

ggplot(AllMerged2_gathered, aes(x = Stats, y = Records, group = Name, colour = Name)) + 
  geom_line(size = 1) + geom_point(color = 'black') + theme_bw() + 
  ggtitle("Five NBA Star Players Stat Comparison") + scale_fill_discrete(name = "Players") +
  theme(axis.text.x = element_text(angle = 45, face = "bold")) + xlab(NULL) + ylab(NULL) + 
  theme(plot.title = element_text(face = "bold"))
```

We can see that James Harden performed the best except the total rebound. As his three points and total points are dominating the other players, he seems to deserve to be the best scorer in this season. Also, compared to the total minutes played, Giannis Antetokounmpo made more scores than the majority did and dominated the total rebound statistic as well. Lastly and impressively, despite his age, Lebron James still performs so well that his statistics are either 2nd or 3rd among the star players.

v) LAL 2020 starting lineup players stat comparison

```{r}
library(fmsb)

LAL_lineups <- nba1920[nba1920$Name == "Alex Caruso" | nba1920$Name == "Anthony Davis" | nba1920$Name == "Danny Green" | 
                  nba1920$Name == "K Caldwell-pope" | nba1920$Name == "Lebron James", 
                  c("TotalMinutesPlayed", "FieldGoalsMade", "FieldGoalsAttempted", "ThreesMade", "ThreesAttempted", "FreeThrowsMade", 
                    "FreeThrowsAttempted", "TotalRebounds", "Assists", "Steals", "Turnovers", "Blocks", "PersonalFouls")]
LAL_lineups$`Field Goal%` <- LAL_lineups$FieldGoalsMade / LAL_lineups$FieldGoalsAttempted * 100
LAL_lineups$`3P%` <- LAL_lineups$ThreesMade / LAL_lineups$ThreesAttempted * 100
LAL_lineups$`Free Thr%` <- LAL_lineups$FreeThrowsMade / LAL_lineups$FreeThrowsAttempted * 100
LAL_lineups <- LAL_lineups[c("TotalMinutesPlayed", "Field Goal%", "3P%", "Free Thr%",
                             "TotalRebounds", "Assists", "Steals", "Turnovers", "Blocks", "PersonalFouls")]
rownames(LAL_lineups) <- c("K Caldwell-pope", "Alex Caruso", "Anthony Davis", "Danny Green", "Lebron James")
colnames(LAL_lineups) <- c("PlayMinutes", "Field Goal%", "3P%", "Free Thr%", "Rebounds", "Assists", "Steals", "Turnovers", "Blocks",
                           "Fouls")
LAL_lineups <- rbind(c(2500, 100, 100, 100, 550, 700, 100, 280, 150, 170), c(1000, 0, 0, 0, 100, 80, 50, 50, 10, 80), LAL_lineups)

par(mfrow = c(2,3))
radarchart(LAL_lineups[1:3, ], title = "K Caldwell-pope (SG)", axistype = 2, pcol=rgb(0.2,0.5,0.5,0.9), pfcol=rgb(0.2,0.5,0.5,0.5),
           plwd=4, cglcol="grey", cglty=1, axislabcol="purple")
radarchart(LAL_lineups[c(1,2,4),], title = "Alex Caruso (PG)", axistype = 2, pcol=rgb(0.8,0.2,0.5,0.9), pfcol=rgb(0.8,0.2,0.5,0.5), 
           plwd=4, cglcol="grey", cglty=1, axislabcol="purple")
radarchart(LAL_lineups[c(1,2,5),], title = "Anthony Davis (PF)", axistype = 2, pcol=rgb(0.7,0.5,0.1,0.9), pfcol=rgb(0.7,0.5,0.1,0.5), 
           plwd=4, cglcol="grey", cglty=1, axislabcol="purple")
radarchart(LAL_lineups[c(1,2,6),], title = "Danny Green (SF)", axistype = 2, pcol=rgb(0.2,0.8,0.4,0.9), pfcol=rgb(0.2,0.8,0.4,0.5), 
           plwd=4, cglcol="grey", cglty=1, axislabcol="purple")
radarchart(LAL_lineups[c(1,2,7),], title = "Lebron James (PG)", axistype = 2, pcol=rgb(0.9,0.4,0.3,0.9), pfcol=rgb(0.9,0.4,0.3,0.5), 
           plwd=4, cglcol="grey", cglty=1, axislabcol="purple")
par(mfrow = c(1,1))

```

The radar charts concern 10 statistics in order to assess each position of LA Lakers starting lineup. 
- Total minutes played
- % of field goals success (field goals made / attempts)
- % of three points success (3P made / attempts)
- % of free throws success (free throws made / attempts)
- Total rebounds
- Assists
- Steals
- Turnovers
- Blocks
- Fouls

vi) Offensive Rebounds vs Defensive Rebounds for Turnovers

```{r}
nba1920$DefensiveRebounds <- nba1920$TotalRebounds - nba1920$OffensiveRebounds  

x <- ggplot(nba1920, aes(x = OffensiveRebounds, y = Turnovers)) + geom_point(color = "dark blue") + 
  theme_bw() + ggtitle("Offensive Rebounds & Turnovers") + xlab("Offensive Rebounds") + 
  theme(plot.title = element_text(face = "bold"))

y <- ggplot(nba1920, aes(x = DefensiveRebounds, y = Turnovers)) + geom_point(color = "orange") + 
  theme_bw() + ggtitle("Defensive Rebounds & Turnovers") + xlab("Defensive Rebounds") + 
  theme(plot.title = element_text(face = "bold"))

grid.arrange(x,y, nrow = 2)
```

There are two kinds of rebounds: offensive and defensive. Offensive rebounds mean that the offensive side recovered the ball so that they keep possessing their turn, while defensive rebounds mean that the defending side gains the possession and the turn changes. Regardless of the types of rebounds, generally the more rebounds the more turnovers. However, if we take a deeper look inside with these two separate rebound types, defensive rebounds have more strong correlation with turnovers than offensive rebounds do. This implies that if the team fails in offensive rebounds, turnovers are more likely to occur than in case the team fails in defensive rebounds.

5) Displaying the home locations of the last 10 champions of the league with GoogleVis

```{r}
library(googleVis)
library(dplyr)
library(tmap)
library(tmaptools)

NBA_Champs <- read_html("https://en.wikipedia.org/wiki/List_of_NBA_champions")
NBA_Champs_data <- NBA_Champs %>% html_nodes("table.wikitable")

Champs <- (NBA_Champs_data[2] %>% html_table())[[1]]
Champs <- tail(Champs, 10) %>% select(Year,`Western champion`,Result,`Eastern champion`)

Champs$`Eastern champion` <- gsub("\\s.*", "", Champs$`Eastern champion`)
Champs$`Western champion` <- gsub("^(\\S*\\s+\\S+).*", "\\1", Champs$`Western champion`)
Champs$`Western champion` <- gsub("Golden State", "San Francisco", Champs$`Western champion`)
Champs$`Western champion` <- gsub("Mavericks", "", Champs$`Western champion`)

Champs$WestWins <- as.numeric(substr(Champs$Result,1,1))
Champs$EastWins <- as.numeric(substr(Champs$Result,3,3))
Champs$Winner <- ifelse(Champs$WestWins > Champs$EastWins, Champs$`Western champion`, Champs$`Eastern champion`)

latlong <- geocode_OSM(unique(Champs$Winner))
latlong <- as_tibble(latlong) %>% rename(Winner = query)

newChamps <- Champs %>% inner_join(latlong)
newChamps$LatLong <- paste(as.character(newChamps$lat), as.character(newChamps$lon), sep = ":")
newChamps$state <- c("Texas", "Florida", "Florida", "Texas", "California", "Ohio", rep("California", 2),
                     "Ontario", "California")

stateChart <- gvisGeoChart(newChamps, "state",
                           options = list(region = "US", resolution = 'provinces'))
cityChart <- gvisGeoChart(newChamps, "LatLong",
                          options = list(region = "US", resolution = 'provinces'))

plot(stateChart)
plot(cityChart)
```

The locations of NBA champions from 2011 to 2020 are presented with googleVis. There are some duplicated areas along with one region in Canada. I plotted two maps (state vs cities) for better understanding. Note that the state map does not indicate Toronto as it’s Canadian region, but the city map shows all the locations by using the longitudes and latitudes.

6) Simple Linear Regression for Total Points by Field Goals Attempted (See comments for the equation)
```{r}
cor(nba1920$FieldGoalsAttempted, nba1920$TotalPoints)  #Field Goals Attempted has a high correlation coefficient with total points

m <- lm(TotalPoints~FieldGoalsAttempted, data = nba1920)
summary(m)    #The equation of the model is y = -14.131 + 1.3x, meaning that each field goal attempt increases 1.3 total points

ggplot(nba1920,aes(x = FieldGoalsAttempted, y = TotalPoints)) + geom_point(color = 'dark blue') +
  geom_smooth(method = "lm", color = 'orange') + theme_bw() + 
  ggtitle("Field Goals Attempted and Total Points") + xlab("Field Goals Attempted") +
  ylab("Total Points") + theme(plot.title = element_text(face = "bold"))
```

Also, below is the interactive plot with ggPredict() for this simple linear regression.
```{r}
library(ggiraph)
library(ggiraphExtra)
library(plyr)
ggPredict(m, se = TRUE, interactive = TRUE)
```

